Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-14 ms.tex
just conclusion left

We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.  We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the low C/O measurement. In one test, we assumed chemical equilibrium but excluded methane opacity. In this test, the C/O value

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. These results $3\,\sigma$

\section{Discussion}

\subsubsection{Condensate Properties}
We also considered physically motivated, self-consistent cloud and haze models based on .  To model WASP-107b's spectrum with self-consistent aerosols, we use the methods described in \citep{fortney08, morley15}. We include models from solar to $50\times$ solar metallicity and solar C/O ratio. We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), varying the cloud sedimentation efficiency from 0.1 to 3. None of the cloudy models are sufficiently low amplitude to match the observed muted signal. We also model an ad hoc photochemical 'soot' layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}. With a sufficiently thick photochemical haze with particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.

@@

\section{Discussion} \label{sec:discuss}
Our data suggest a sub-solar methane abundance at FIXME $\sigma$ confidence. This result was surprising because methane is expected to be the dominant carbon-bearing molecule at WASP-107b's temperature \citep{moses13}, but super-solar metal

@@

low C/O ratio - planetesimal enrichment (cite Nestor)
metallicity 30x solar - how does this compare to mass/atmospheric metallicity 
clouds

\subsection{Comparative planetology with HAT-P-11b, HAT-P-26b}

\section{Conclusions and Future Plans}
WASP-107b is slated to be observed with numerous facilities. Transits were recently observed with the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels. Spitzer eclipses are planned (Program FIXME; PI L. Kreidberg). In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.

>>>>

We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.  We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O. In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$. We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.  Assuming  solar C/O and $3\,\times$ solar metallicity (the median retrieved value for the chemical equilbrium), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.  Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.

\subsection{Condensate Properties}
In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.  We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar. We vary the cloud sedimentation efficiency from 0.1 to 3 (highly lofted, small particulate clouds). None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.

We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}. With the most efficient haze production and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations. However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.  %Sulfur-based hazes are another possibility \citep{gao17}.
Further modeling of the condensates is needed for a satisfactory explanation of the data.

@@

\section{Discussion and Conclusions} \label{sec:discuss}
We combined our high precision transmission spectrum of WASP-107b with interior structure modeling to assess the planet's atmospheric composition.

Upper limit on metallicity. The planet's radius is so large that we put an upper limit on the metallicity (additional metals would compress it).  metallicity 30x solar - how does this compare to mass/atmospheric metallicity

Methane depletion
low C/O ratio - planetesimal enrichment (cite Nestor)
internal heat, directly imaged planets

Thick, high-altitude condensate layer

@@

WASP-107b is slated to be observed with numerous facilities. Transits were recently observed with the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels. Spitzer eclipses are planned (Program FIXME; PI L. Kreidberg). In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.
